---
title: "Water Quality Presentation 1"
author: "Group 3: Stephanie Fissel, Beza Gashe, Claire Yoon"
date: "January 3, 2023"
output: 
  html_document:
    theme: lumen
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
editor_options: 
chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(aod)
library(vtable)
```

# Problem
...

# Background
...

# Question
...

# Data Source

### ***[Water Quality](https://www.kaggle.com/datasets/adityakadiwal/water-potability?select=water_potability.csv)***
A dataset containing water quality metrics for 3276 different water bodies. <br/>
Key metrics include: <br/>
    
* pH value
    + indicator of acidic or alkaline condition of water status
* Hardness
    + caused by calcium and magnesium salts
* Solids (Total dissolved solids - TDS)
   + indicator of level of mineralization of water
* Chloramines
   + level of disinfectant in water
* Sulfates
   + naturally occurring substance present in minerals, soil, and rocks
* Conductivity
   + Item 2a
* Organic carbon
   + measure of the total amount of carbon in organic compounds in pure water
* Trihalomethanes
   + chemicals found in water treated with chlorine
* Turbidity
   + depends on quantity of solid matter pressent in suspended state

Target metric: <br/>

* Potability
   + indicator of whether water is safe for human consumption where 1=Potable 
   and 0=Not Potable


#### Read in the data
```{r}
water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")

head(water_quality)
```

### Clean the data
#### Eliminate the "Sulfate" column: 
more info why...

```{r}
water_quality <- subset(water_quality, select = -c(Sulfate))
head(water_quality)
```
#### Remove rows containing at least one NA
```{r}
water_quality <- na.omit(water_quality)
```
#### Converting 'Potability' to a factor (categorical variable)
```{r}
water_quality$Potability <- factor(water_quality$Potability)
head(water_quality)
```

# Descriptive Statistics {.tabset}
## **SUMMARY STATS** {.tabset}

```{r}
st(water_quality)
```




## **DISTRIBUTIONS** {.tabset .tabset-fade .tabset-pills}

### pH
```{r}
# Histogram
hist(water_quality$ph, prob = TRUE,
     col = "white",
     main = "pH Level Distribution",
     xlab = "pH Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$ph, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0, 0, alpha = 0.5))

# Box around the plots
box()
```

### Hardness
```{r}
# Histogram
hist(water_quality$Hardness, prob = TRUE,
     col = "white",
     main = "Hardness Level Distribution",
     xlab = "Hardness Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Hardness, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0.4, 0, alpha = 0.5))

# Box around the plots
box()
```

### Solids
```{r}
# Histogram
hist(water_quality$Solids, prob = TRUE,
     col = "white",
     main = "Solids Level Distribution",
     xlab = "Solids Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Solids, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 1, 0, alpha = 0.5))

# Box around the plots
box()
```

### Chloramines
```{r}
# Histogram
hist(water_quality$Chloramines, prob = TRUE,
     col = "white",
     main = "Chloramines Level Distribution",
     xlab = "Chloramines Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Chloramines, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.4, 0, alpha = 0.5))

# Box around the plots
box()
```

### Conductivity
```{r}
# Histogram
hist(water_quality$Conductivity, prob = TRUE,
     col = "white",
     main = "Conductivity Level Distribution",
     xlab = "Conductivity Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Conductivity, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.6, 1, alpha = 0.5))

# Box around the plots
box()
```

### Organic Carbon
```{r}
# Histogram
hist(water_quality$Organic_carbon, prob = TRUE,
     col = "white",
     main = "Organic_carbon Level Distribution",
     xlab = "Organic_carbon Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Organic_carbon, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.2, 1, alpha = 0.5))

# Box around the plots
box()
```

### Trihalomethanes
```{r}
# Histogram
hist(water_quality$Trihalomethanes, prob = TRUE,
     col = "white",
     main = "Trihalomethanes Level Distribution",
     xlab = "Trihalomethanes Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Trihalomethanes, horizontal = TRUE, axes = FALSE,
        col = rgb(0.6, 0, 1, alpha = 0.5))

# Box around the plots
box()
```

### Turbidity
```{r}
# Histogram
hist(water_quality$Turbidity, prob = TRUE,
     col = "white",
     main = "Turbidity Level Distribution",
     xlab = "Turbidity Level",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Turbidity, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0.2, 0.8, alpha = 0.5))

# Box around the plots
box()
```






