---
title: "Water Quality Presentation 2"
author: "Group 3: Stephanie Fissel, Beza Gashe, Claire Yoon"
date: "January 4, 2023"
output: 
  html_document:
    theme: lumen
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
editor_options: 
chunk_output_type: inline
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(aod)
library(vtable)
library(scales)
library(reshape2)
library(gridExtra)
library(grid)
library(cowplot)
library(dplyr)
library(data.table)
library(e1071)
library(htmltools)
library(devtools)
library(caret)
library(NbClust)
library(randomForest)
library(rio)
library(plyr)
library(ROCR)
library(rpart)
library(psych)
library(pROC)
library(rpart.plot)
library(rattle)
library(class)
library(MLmetrics)
library(mltools)
library(RColorBrewer)
```

# Updates
* Working on  ML models
    + Random Forest & Decision Tree
* Working on Shiny app using R
    + Visualizations will help us see trends in the data that may be explained by
    our ML models and further understand what levels of substances are contained 
    in drinking water that is claimed to be "potable" though they can have
    adverse effects
    
# Random Forest Classifier

### Load the Data
```{r}
getwd()
#water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")
water_quality <- read.csv("~/Desktop/R/water-quality-project/water_potability.csv")
```


### Finish any other data prep (one-hot encode, reduce factor levels)
```{r}
dim(water_quality) #10 columns, 3276 rows
sum(is.na(water_quality)) #1434 missing values
colSums(is.na(water_quality))#?colSums
#missing values found in ph, sulfate, and trihalomethanes
#Delete missing values
water_quality <- water_quality[complete.cases(water_quality),]
sum(is.na(water_quality))
#str(water_quality)
#converting target variable, potability to a factor:
water_quality$Potability <- factor(water_quality$Potability)
#str(water_quality)
```

No need to standardize the data

### Prevalence
```{r}
water_quality1<-names(select_if(water_quality, is.factor))
water_quality1
#str(water_quality)
table(water_quality$Potability)[2]/sum(table(water_quality$Potability))
#prevalence=0.40
water_quality$Potability <- recode(water_quality$Potability, 
                           '0' = 'Non_Potable', 
                            '1' = 'Potable')
```


### Test, tune and training sets 
```{r}
set.seed(1999)
part_index_1 <- caret::createDataPartition(water_quality$Potability,
                                           times=1,
                                           p = 0.80,
                                           groups=1,
                                           list=FALSE)
train <- water_quality[part_index_1, ]
tune_and_test <- water_quality[-part_index_1, ]
train
tune_and_test_index <- createDataPartition(water_quality$Potability,
                                           p = .1,
                                           list = FALSE,
                                           times = 1)
tune <- tune_and_test[tune_and_test_index, ]
test <- tune_and_test[-tune_and_test_index, ]
dim(train)
#1609  14
dim(test)
#367   14
dim(tune)
#202  14
```


### Calculate the initial mtry level 
```{r}
water_quality_tune <- function(x){
  water_quality <- dim(x)[2]-1
  sqrt(water_quality)
}
water_quality_tune(water_quality)
# = 3
```

### Run the initial RF model with 1000 trees
```{r}
water_quality_RF = randomForest(as.factor(Potability)~.,          
                            train,     
                            ntree = 1000,  
                            mtry = 3, 
                            replace = TRUE,      
                            sampsize = 100,      
                            nodesize = 4,        
                            
                            importance = TRUE,   
                           
                            proximity = FALSE, 
                            norm.votes = TRUE,
                            do.trace = TRUE,     
                            keep.forest = TRUE, 
                            keep.inbag = TRUE)  
view(water_quality_RF)
```

### Variable Importance
```{r}
set.seed(1)
varImp(water_quality_RF) 
```

### Using the training and tune datasets to tune the model 
#### - In consideration of the number of trees, the number of variables to sample and the sample size that optimize the model output.
```{r}
#### Optimize the random forest model ####
set.seed(1984)	
water_quality_RF_2 = randomForest(as.factor(Potability)~.,          #<- 
                              train,    
                              ntree = 1000,     
                              mtry = 6,          
                              replace = TRUE,   
                             
                              sampsize = 200,    
                              nodesize = 10, 
                              importance = TRUE,  
                              proximity = FALSE, 
                              norm.votes = TRUE, 
                              do.trace = TRUE, 
                              keep.forest = TRUE, 
                              keep.inbag = TRUE) 
```

### Evaluate the model using the test dataset. 
```{r}
water_quality_predict = predict(water_quality_RF_2,      #<- a randomForest model
                            test,      #<- the test data set to use
                            type = "response",   #<- what results to produce, see the help menu for the options
                            predict.all = TRUE)
str(water_quality_predict)
water_quality_RF_2$confusion
#            Non_Potable Potable class.error
#Non_Potable         887      73  0.07604167
#Potable             464     185  0.71494607
water_quality_RF$confusion
#            Non_Potable Potable class.error
#Non_Potable         910      50  0.05208333
#Potable             525     124  0.80893683
library(caret)
confusionMatrix(as.factor(water_quality_predict$aggregate),as.factor(test$Potability),positive = "Potable", 
                dnn=c("Prediction", "Actual"), mode = "everything")
```

# Decision Tree
```{r}
#water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")
#water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")
water_quality <- read.csv("~/Desktop/R/water-quality-project/water_potability.csv")
#head(water_quality)
#dim(water_quality)
```

### Clean the data
There were 1/3 missing values in Sulfate variable, but we didn’t delete it because sulfate showed one of the crucial factors for predicting non-potability because it leads to diarrhea and some problems. After handling with the missing values, we have 2011 objects.
```{r}
water_quality <- water_quality[complete.cases(water_quality), ]
#dim(water_quality)
```

### Rename and factorize the 'Potability' variable
We renamed the categorical values (1 and 0) to "potable" and "non_potable" </br>
We are focusing on non potability of drinking water, so our positive value is "non_potable"
```{r}
water_quality$Potability <- as.factor(water_quality$Potability)
water_quality$Potability <- fct_collapse(water_quality$Potability, 
                           potable = "1",
                           non_potable = "0")
```

### Splitting the data with tuning and test & Creating Decision Trees
#### First Decision Tree Model
```{r}
#We do not need to check for correlated variables because correlation does not impact decision trees.
#Decision trees make greedy, localized decisions that are not dependent on previous steps or other variables in the tree model.  
set.seed(777)
partition <- caret::createDataPartition(water_quality$Potability,
                                           times=1,
                                           p = 0.80,
                                           groups=1,
                                           list=FALSE)
train <- water_quality[partition, ]
tune_and_test <- water_quality[-partition, ]
#train
tune_and_test_index <- createDataPartition(tune_and_test$Potability,
                                           p = .5,
                                           list = FALSE,
                                           times = 1)
tune <- tune_and_test[tune_and_test_index, ]
test <- tune_and_test[-tune_and_test_index, ]
#dim(train)
#dim(test) 
#dim(tune)
# Build the model
# Choose the features and classes, slightly different approach for caret, need to create features and target sets from the training data.
#str(water_quality)
features <- train[,-10] # dropping the target variable
target <- train$Potability
#str(features)
#str(target)
fitControl <- trainControl(method = "repeatedcv",
                          number = 10,
                          repeats = 5,
                          classProbs = TRUE,
                          summaryFunction = twoClassSummary) 
tree.grid <- expand.grid(maxdepth=c(1:6))
# Train the model without setting tree grid
set.seed(777)
water_mdl1 <- train(x=features,
                y=target,
                method="rpart2",
                trControl=fitControl,
                metric="ROC")
plot(water_mdl1)
varImp(water_mdl1)
rpart.plot(water_mdl1$finalModel, type=1,extra=101)
water_mdl1 # maxdepth is 11
```


#### Second Decision Tree Model
```{r}
# Train the model
set.seed(777)
water_mdl2 <- train(x=features,
                y=target,
                method="rpart2",
                trControl=fitControl,
                tuneGrid=tree.grid,
                metric="ROC")
plot(water_mdl2)
varImp(water_mdl2)
rpart.plot(water_mdl2$finalModel, type=1,extra=101)
water_mdl2 # maxdepth is 6
```


### Confusion Matrix, Statistics, and Density Plot on Two Decision Trees
#### First Model

```{r}
predictandCM<- function(model,data,modeltype,ref)
{
  pred <-predict(model,data,type=modeltype)
  confusionMatrix(pred, reference=ref, positive = 'non_potable')
}
predictandCM(water_mdl1, tune, "raw", tune$Potability)
water_pred_tune = predict(water_mdl1,tune,tune$Potability, type= "prob")
water_pred_tune_labels = predict(water_mdl1,tune,tune$Potability,type = "raw")
plot(density(water_pred_tune$non_potable))
```

#### Second Model
```{r}
predictandCM(water_mdl2, tune, "raw", tune$Potability) 
water_pred_tune2 = predict(water_mdl2,tune,tune$Potability, type= "prob")
water_pred_tune_labels2 = predict(water_mdl2,tune,tune$Potability,type = "raw")
plot(density(water_pred_tune2$non_potable))
#View(as_tibble(water_pred_tune_labels))
```


### Understanding Confusion Matrix
True Positive(TP) </br>
- Predicted condition: non-potable </br>
- Actual condition: non-potable </br>
</br>
False Positive(FP) </br>
- Predicted condition: non-potable </br>
- Actual condition: potable </br>
</br>
False Negative(FN) </br>
- Predicted condition: potable </br>
- Actual condition: non-potable </br>
</br>
True Negative(TN) </br>
- Predicted condition: potable </br>
- Actual condition: potable </br>
<br>
Considering the real world, False Positive doesn't really make any problems because people do not drink water if it is predicted as non-potable even though it is potable. However, the opposite case (False Negative) will cause significant problems (people will be sick), so we need to make sure to reduce False Negative portion and the model should be reliable.

```{r}
# Use the the confusion matrix function on your predictions to check a variety of metrics and comment on the metric that might be best for this type of analysis given your question.  
water_eval <- caret::confusionMatrix(water_pred_tune_labels2, 
                as.factor(tune$Potability), 
                dnn=c("Prediction", "Actual"),
                positive="non_potable",
                mode = "everything")
#water_eval
```

### Adjusting Threshold with the Second Decision Tree Model
We were adjusting threshold between 0.2 and 0.8, then found 0.6 is a good point considering the balance of sensitivity and specificity. However, the accuracy of the model is still low, so we need to improve on our model and reduce the False Negative values.
```{r}
# With the percentages you generated in step 10,select several different threshold levels using the threshold function we created and interpret the results. What patterns do you notice, did the evaluation metrics change?
adjust_thres <- function(x, y, z) {
  #x=pred_probablities, y=threshold, z=tune_outcome
  thres <- as.factor(ifelse(x > y, "non_potable","potable"))
  confusionMatrix(thres, z, positive = "non_potable", dnn=c("Prediction", "Actual"), mode = "everything")
}
adjust_thres(water_pred_tune2$non_potable,y=.6,tune$Potability)
```

# ML Models
* Next steps:
    + Improving the model and accuracy

# Shiny App (in progress)
* Next steps:
    + Deploy to Shiny.io
    + Add more context
    + Possibly add interactive inputs
    + Analyze visualizations and summarize conclusions


```{r}
#water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")
water_quality <- read.csv("~/Desktop/R/water-quality-project/water_potability.csv")

water_quality <- na.omit(water_quality)
water_quality <- na.omit(water_quality)
water_quality$Potability <- factor(water_quality$Potability)

np <- round(1998/3276 * 100)
p <- round(1278/3276 * 100)
# Subset test data
data <- data.frame(
  category=c("Non-Potable", "Potable"),
  count=c(np, p)
)

# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, ": ", data$count, "%")



shinyApp(
  ui <- fluidPage(
    # App title
    titlePanel("METRICS"),
    mainPanel(
      
      tabsetPanel(type = "tabs",
                  tabPanel("Overall", plotOutput("overall"), width = "100%"),
                  tabPanel("pH", plotOutput("ph")),
                  tabPanel("Hardness", plotOutput("hardness")),
                  tabPanel("Solids", plotOutput("solids")),
                  tabPanel("Chloramines", plotOutput("chloramines")),
                  tabPanel("Conductivity", plotOutput("conductivity")),
                  tabPanel("Organic Carbon", plotOutput("organic_carbon")),
                  tabPanel("Trihalomethanes", plotOutput("trihalomethanes")),
                  tabPanel("Turbidity", plotOutput("turbidity")),
                  ))
    ),
  
  server <- function(input, output) {
    
    output$overall <- renderPlot({
      
      ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
        geom_rect() +
        geom_label( x=3.5, aes(y=labelPosition, label=label), size=5) +
        scale_fill_manual(values = c("#E86363", "#63E871")) + 
        coord_polar(theta="y") +
        xlim(c(2, 4)) +
        theme_void() +
        theme(legend.position = "none")
    }, height = 700, width = 900)
    
    output$ph <- renderPlot({
      
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=ph)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=ph)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = ph, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_x_continuous(breaks= 0:14) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 6.5, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 8.5, color = "darkslategrey", linetype = "dashed", size=0.3) +
        labs(x = "pH Level (pH)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 5, y = 50, label = "< 6.5 is acidic") +
        annotate("text", x = 10, y = 50, label = "> 8.5 is basic") + 
        annotate("text", x = 7.5, y = 62, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 7.5, y = 60, label = "levels", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("pH Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
      })
      
    output$hardness <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Hardness)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Hardness)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Hardness, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 75, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 150, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 300, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Hardness Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 75, y = 45, label = "< 75 is soft") +
        annotate("text", x = 110, y = 50, label = "> 75 is moderately hard") +
        annotate("text", x = 170, y = 50, label = "> 150 is hard") +
        annotate("text", x = 305, y = 50, label = "> 300 is very hard")
      
      title <- ggdraw() + draw_label("Hardness Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$solids <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Solids)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Solids)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Solids, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 500, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 1000, color = "darkslategrey", linetype = "dashed", size=0.3) +
        labs(x = "Solid Level (mg/l)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 6000, y = 50, label = "> 1000 is high") + 
        annotate("text", x = 800, y = 62, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 800, y = 60, label = "levels", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Solid Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$chloramines <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Chloramines)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Chloramines)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Chloramines, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 4, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Chloramine Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 4.8, y = 45, label = "> 4 is high") +
        annotate("text", x = 4, y = 55, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 4, y = 53, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Chloramine Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$conductivity <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Conductivity)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(Conductivity)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Conductivity, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 400, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Conductivity Level (μS/cm)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 460, y = 42, label = "> 400 is high") +
        annotate("text", x = 400, y = 47, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 400, y = 45, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Conductivity Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$organic_carbon <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Organic_carbon)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Organic_carbon)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Organic_carbon, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 2, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Organic Carbon Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 4, y = 35, label = "> 2 is high") + 
        annotate("text", x = 2.5, y = 43, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 2.5, y = 41, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Organic Carbon Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$trihalomethanes <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Trihalomethanes)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Trihalomethanes)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Trihalomethanes, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 80, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Trihalomethanes Level (ppm)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 90, y = 40, label = "> 80 is high") +
        annotate("text", x = 80, y = 47, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 80, y = 45, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Trihalomethanes Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$turbidity <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Turbidity)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Turbidity)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Turbidity, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 5, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Turbidity Level (NTU)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 5.4, y = 37, label = "> 5 is high") +
        annotate("text", x = 5, y = 45, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 5, y = 43, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Turbidity Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    })
```






