---
title: "Water Quality Presentation 1"
author: "Group 3: Stephanie Fissel, Beza Gashe, Claire Yoon"
date: "January 3, 2023"
output: 
  html_document:
    theme: lumen
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
editor_options: 
chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(aod)
library(vtable)
library(scales)
```

# Problem
Access to safe drinking water is essential to health and a basic human right, 
though many developing countries and even parts of the U.S. like Flint, Michigan are living with 
contaminated drinking water. With the increased prevalence of climate change, 
drinking water is becoming even more negatively impacted by contaminants. In 
these instances, non-potable drinking water containing unsafe levels of 
contaminants can have serious adverse health effects, such as gastrointestinal 
illnesses, cancer, and transmissible diseases. 

# Question
There are various water quality metrics that may contribute to the potability of
drinking water. We want to analyze and determine what metrics may more heavily 
contribute to water being non-potable which could, in the real world, help 
formulate effective policy for health protection through implementation of 
systems that specifically de-contaminate the water of these substances. In some 
regions, it has been shown that investments in water supply and sanitation can yield a new 
economic benefit, since the reductions in adverse health effects and health care 
costs outweigh the costs of undertaking the interventions. Continuing with this 
idea, we can then answer the question, "How well can a Machine Learning model 
accurately predict the non-potability of water?"

# Data Source

### ***[Water Quality](https://www.kaggle.com/datasets/adityakadiwal/water-potability?select=water_potability.csv)***
A dataset containing water quality metrics for 3276 different water bodies. <br/>
Key metrics include: <br/>
    
* pH value
    + acidic or alkaline condition
* Hardness
    + caused by calcium and magnesium salts
* Solids (Total dissolved solids - TDS)
   + mineralization of organic minerals or salts
* Chloramines
   + disinfectant (chlorine + ammonia = chloramines)
* Sulfates
   + naturally occurring substance present in minerals, soil, and rocks
* Conductivity
   + determined by amount of dissolved solids
* Organic carbon
   + carbon in decaying natural organic matter and synthetic sources
* Trihalomethanes
   + chemicals found in water treated with chlorine
* Turbidity
   + depends on quantity of solid matter present; measure of light emitting 
   properties of water

Target metric: <br/>

* Potability
   + indicator of whether water is safe for human consumption where 1 = Potable 
   and 0 = Not Potable


#### Read in the data
```{r}
water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")

head(water_quality)
```

### Clean the data
```
#### Remove rows containing at least one NA
```{r}
water_quality <- na.omit(water_quality)
```
#### Converting 'Potability' to a factor (categorical variable)
```{r}
water_quality$Potability <- factor(water_quality$Potability)
```

# Descriptive Statistics {.tabset}
## **SUMMARY STATS** {.tabset}

```{r}
st(water_quality)
```

## **POTABLE VS. NON-POTABLE** {.tabset}
```{r}
# non-potable
# colSums(water_quality==0) #--> 1998
# potable
# colSums(water_quality==1) #--> 1278
np <- round(1998/3276 * 100)
p <- round(1278/3276 * 100)
# Subset test data
data <- data.frame(
  category=c("Non-Potable", "Potable"),
  count=c(np, p)
)

# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, ": ", data$count, "%")

# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=3) +
  scale_fill_brewer(palette=1) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")
```

## **DISTRIBUTIONS** {.tabset .tabset-fade .tabset-pills}

### pH
**Recommended Level:** 6.5-8.5 pH (neutral on pH scale)
```{r}
# Histogram
hist(water_quality$ph, prob = TRUE,
     col = "white",
     main = "pH Level Distribution",
     xlab = "pH Level (pH)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$ph, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0, 0, alpha = 0.5))

# Box around the plots
box()
```

### Hardness
**Recommended Level:** None <br/>
Hardness levels > 150 mg/L are considered hard water. If water is hard, it can 
reduce efficiency of water 
treatment.
```{r}
# Histogram
hist(water_quality$Hardness, prob = TRUE,
     col = "white",
     main = "Hardness Level Distribution",
     xlab = "Hardness Level (mg/L)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Hardness, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0.4, 0, alpha = 0.5))

# Box around the plots
box()
```

### Solids
**Recommended Level:** 500-1000 mg/l
```{r}
# Histogram
hist(water_quality$Solids, prob = TRUE,
     col = "white",
     main = "Solids Level Distribution",
     xlab = "Solids Level (mg/l)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Solids, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 1, 0, alpha = 0.5))

# Box around the plots
box()
```

### Chloramines
**Recommended Level:** 4 mg/L or less
```{r}
# Histogram
hist(water_quality$Chloramines, prob = TRUE,
     col = "white",
     main = "Chloramines Level Distribution",
     xlab = "Chloramines Level (mg/L)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Chloramines, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.4, 0, alpha = 0.5))

# Box around the plots
box()
```

### Conductivity
**Recommended Level:** 400 μS/cm or less
```{r}
# Histogram
hist(water_quality$Conductivity, prob = TRUE,
     col = "white",
     main = "Conductivity Level Distribution",
     xlab = "Conductivity Level (μS/cm)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Conductivity, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.6, 1, alpha = 0.5))

# Box around the plots
box()
```

### Organic Carbon
**Recommended Level:** 2 mg/L or less
```{r}
# Histogram
hist(water_quality$Organic_carbon, prob = TRUE,
     col = "white",
     main = "Organic Carbon Level Distribution",
     xlab = "Organic Carbon Level (mg/L)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Organic_carbon, horizontal = TRUE, axes = FALSE,
        col = rgb(0, 0.2, 1, alpha = 0.5))

# Box around the plots
box()
```

### Trihalomethanes
**Recommended Level:** 80 ppm or less
```{r}
# Histogram
hist(water_quality$Trihalomethanes, prob = TRUE,
     col = "white",
     main = "Trihalomethanes Level Distribution",
     xlab = "Trihalomethanes Level (ppm)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Trihalomethanes, horizontal = TRUE, axes = FALSE,
        col = rgb(0.6, 0, 1, alpha = 0.5))

# Box around the plots
box()
```

### Turbidity
**Recommended Level:** 5 NTU or less
```{r}
# Histogram
hist(water_quality$Turbidity, prob = TRUE,
     col = "white",
     main = "Turbidity Level Distribution",
     xlab = "Turbidity Level (NTU)",
     ylab = "Count")

# Add new plot
par(new = TRUE)

# Box plot
boxplot(water_quality$Turbidity, horizontal = TRUE, axes = FALSE,
        col = rgb(1, 0.2, 0.8, alpha = 0.5))

# Box around the plots
box()
```






