---
title: "Water Quality Presentation 2"
author: "Group 3: Stephanie Fissel, Beza Gashe, Claire Yoon"
date: "January 4, 2023"
output: 
  html_document:
    theme: lumen
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
editor_options: 
chunk_output_type: inline
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(aod)
library(vtable)
library(scales)
library(reshape2)
library(gridExtra)
library(grid)
library(cowplot)
library(dplyr)
library(data.table)
```

# Updates
* Working on  ML models
    + ...
* Working on Shiny app using R
    + Visualizations will help us see trends in the data that may be explained by
    our ML models and further understand what levels of substances are contained 
    in drinking water that is claimed to be "potable" though they can have 
    adverse effects
    
# ML Models
* Next steps:
    + ...
    
```{r}
# (CODE HERE)
```



# Shiny App (in progress)
* Next steps:
    + Deploy to Shiny.io
    + Add more context
    + Possibly add interactive inputs
    + Analyze visualizations and summarize conclusions
```{r}
water_quality <- read.csv("/Users/stephaniefissel/Library/Mobile Documents/com~apple~CloudDocs/ds final project/water_potability.csv")

water_quality <- na.omit(water_quality)
water_quality <- na.omit(water_quality)
water_quality$Potability <- factor(water_quality$Potability)

np <- round(1998/3276 * 100)
p <- round(1278/3276 * 100)
# Subset test data
data <- data.frame(
  category=c("Non-Potable", "Potable"),
  count=c(np, p)
)

# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, ": ", data$count, "%")


shinyApp(
  ui <- fluidPage(
    # App title
    titlePanel("METRICS"),
    mainPanel(
      
      tabsetPanel(type = "tabs",
                  tabPanel("Overall", plotOutput("overall"), width = "100%"),
                  tabPanel("pH", plotOutput("ph")),
                  tabPanel("Hardness", plotOutput("hardness")),
                  tabPanel("Solids", plotOutput("solids")),
                  tabPanel("Chloramines", plotOutput("chloramines")),
                  tabPanel("Conductivity", plotOutput("conductivity")),
                  tabPanel("Organic Carbon", plotOutput("organic_carbon")),
                  tabPanel("Trihalomethanes", plotOutput("trihalomethanes")),
                  tabPanel("Turbidity", plotOutput("turbidity")),
                  ))
    ),
  
  server <- function(input, output) {
    
    output$overall <- renderPlot({
      
      ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
        geom_rect() +
        geom_label( x=3.5, aes(y=labelPosition, label=label), size=5) +
        scale_fill_manual(values = c("#E86363", "#63E871")) + 
        coord_polar(theta="y") +
        xlim(c(2, 4)) +
        theme_void() +
        theme(legend.position = "none")
    }, height = 700, width = 900)
    
    output$ph <- renderPlot({
      
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=ph)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=ph)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = ph, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_x_continuous(breaks= 0:14) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 6.5, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 8.5, color = "darkslategrey", linetype = "dashed", size=0.3) +
        labs(x = "pH Level (pH)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 5, y = 50, label = "< 6.5 is acidic") +
        annotate("text", x = 10, y = 50, label = "> 8.5 is basic") + 
        annotate("text", x = 7.5, y = 62, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 7.5, y = 60, label = "levels", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("pH Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
      })
      
    output$hardness <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Hardness)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Hardness)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Hardness, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 75, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 150, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 300, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Hardness Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 75, y = 45, label = "< 75 is soft") +
        annotate("text", x = 110, y = 50, label = "> 75 is moderately hard") +
        annotate("text", x = 170, y = 50, label = "> 150 is hard") +
        annotate("text", x = 305, y = 50, label = "> 300 is very hard")
      
      title <- ggdraw() + draw_label("Hardness Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$solids <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Solids)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Solids)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Solids, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 500, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        geom_vline(xintercept = 1000, color = "darkslategrey", linetype = "dashed", size=0.3) +
        labs(x = "Solid Level (mg/l)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 6000, y = 50, label = "> 1000 is high") + 
        annotate("text", x = 800, y = 62, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 800, y = 60, label = "levels", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Solid Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$chloramines <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Chloramines)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Chloramines)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Chloramines, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 4, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Chloramine Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 4.8, y = 45, label = "> 4 is high") +
        annotate("text", x = 4, y = 55, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 4, y = 53, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Chloramine Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$conductivity <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Conductivity)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(Conductivity)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Conductivity, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 400, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Conductivity Level (μS/cm)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 460, y = 42, label = "> 400 is high") +
        annotate("text", x = 400, y = 47, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 400, y = 45, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Conductivity Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$organic_carbon <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Organic_carbon)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Organic_carbon)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Organic_carbon, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 2, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Organic Carbon Level (mg/L)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 4, y = 35, label = "> 2 is high") + 
        annotate("text", x = 2.5, y = 43, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 2.5, y = 41, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Organic Carbon Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$trihalomethanes <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Trihalomethanes)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Trihalomethanes)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Trihalomethanes, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 80, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Trihalomethanes Level (ppm)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 90, y = 40, label = "> 80 is high") +
        annotate("text", x = 80, y = 47, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 80, y = 45, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Trihalomethanes Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    output$turbidity <- renderPlot({
      nf <- layout( matrix(c(1,2), ncol=1))
      par(mar=c(0, 3.1, 1.1, 2.1))
      
      p1_1 <- ggplot(subset(water_quality, Potability %in% c("0")), aes(x=Turbidity)) +
        geom_boxplot(fill="#E86363", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      p1_2 <- ggplot(subset(water_quality, Potability %in% c("1")), aes(x=Turbidity)) +
        geom_boxplot(fill="#63E871", alpha=0.7) +
        labs(x = " ") +
        theme_bw() +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), axis.text.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())
      
      par(mar=c(4, 3.1, 1.1, 2.1))
      
      p2 <- ggplot(water_quality, aes(x = Turbidity, color = Potability, fill=Potability)) + 
        geom_histogram(position= "identity", alpha = 0.4, bins=100) +
        scale_color_manual(values=c("#E86363", "#63E871"), guide = "none") +
        scale_fill_manual(values=c("#E86363", "#63E871"), labels = c("Non-Potable", "Potable")) +
        geom_vline(xintercept = 5, color = "darkslategrey", linetype = "dashed", size=0.3) + 
        labs(x = "Turbidity Level (NTU)", y = "Count") + 
        theme(text=element_text(size=15, family = "Helvetica"), legend.position=c(0.85, 0.5)) + 
        annotate("text", x = 5.4, y = 37, label = "> 5 is high") +
        annotate("text", x = 5, y = 45, label = "recommended", size=3.2, color="darkslategrey") +
        annotate("text", x = 5, y = 43, label = "threshold", size=3.2, color="darkslategrey")
      
      title <- ggdraw() + draw_label("Turbidity Level Distribution", fontface='bold', size=20)
      plot_grid(title, p1_1, p1_2, p2, ncol=1, align="v", rel_heights = c(0.6, 0.8, 0.8, 6))
    })
    
    })
```






